Current Cropland, Production, and Yield

### 2016-01-19 Rmd/biodiversity-layer.Rmd
Created new carbon layers
Updated cropland datasets, calculating current productivity for maize and soybean from FAO and USDA statistics.

# Sum for total - the inverse of this will be the farmable areas
remfrac <- calc(s, sum)
remfrac <- mask(remfrac, mgrid)

# Calculate intactness
w <- matrix(nrow = 11, ncol = 11)
w[] <- 1
intact <- focal(remfrac, w = w, fun = mean, na.rm = TRUE)
# intact <- mask(intact, mgrid)
# plot(intact)

# Calculate a veg type weight based on relative remaining area
veg_dt <- dtraster::as.data.table(sm)  # stack and convert to dt
veg_dt <- na.omit(veg_dt)
vnms <- names(veg_dt)[-c(1:2)]
vtots <- veg_dt[, lapply(.SD, function(x) sum(x * 100)), 
                .SDcols = vnms]  
wgti <- round(1 / (vtots / nrow(veg_dt)), 2)  # bd weights (1 = rarest)
# wgt <- round(vtots / nrow(veg_dt), 4)  # biodiversity weights (1 = most common)
# plot(unlist(wgti), unlist(wgt))

@@

# plots to examine them
labs <- gsub("veg", "", names(wgtsn))
# par(mfrow = c(2, 1), mar = c(2, 2, 2, 2))
# weights against protected proportion (NPs + GMAs)
plot(wgtsn, prop_prot2, pch = "", ylab = "")
text(wgtsn, prop_prot2, labels = labs, cex = 0.75)
# weights against protected proportion (NPs only)
plot(wgtsn, prop_prot1, pch = "", ylab = "")
text(wgtsn, prop_prot1, labels = labs, cex = 0.75)
# additive weights factoring in protected proportion versus protected proportion
plot(wgts_prop2, prop_prot, pch = "", ylab = "")
text(wgts_prop2,  prop_prot, labels = labs, cex = 0.75)
# BD + NP protected proportion versus NP protected proportion
plot(wgts_prop_np, prop_prot1, pch = "", ylab = "")
text(wgts_prop_np,  prop_prot1, labels = labs, cex = 0.75)
# protected proportion weighted versus same but log-transformed
plot(wgts_prop, wgts_propl, pch = "", ylab = "")
text(wgts_prop,  wgts_propl, labels = labs, cex = 0.75)
# Bd weights against BD + protected proportion weights
plot(wgtsn, wgts_prop, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn, wgts_prop, labels = labs, cex = 0.75)
# Bd weights against BD + NP protected proportion weights
plot(wgtsn, wgts_prop_np, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn, wgts_prop_np, labels = labs, cex = 0.75)
# Bd weights versus log-transformed BD + protected proportion weights
plot(wgtsn, wgts_propl, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn,  wgts_propl, labels = labs, cex = 0.75)
# The same but against multiplicative weights
plot(wgtsn, wgtsn * unlist(1 - prop_prot), pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn,  wgtsn * unlist(1 - prop_prot), labels = labs, cex = 0.75)
# sort(wgts_prop); sort(wgts_prop_np)
# sort(wgtsn)

@@

# final biodiversity layer
bdf <- calc(stack(bdr, pafix), max)
plot(remfrac)
plot(bdf)

writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"))

>>>>

# keep for assessing impacts
writeRaster(sm, filename = full_path(p_bd, "ZA-vegremain.tif"), overwrite = TRUE)

# Sum for total - the inverse of this will be the farmable areas
remfrac <- calc(s, sum)
remfrac <- mask(remfrac, mgrid)
writeRaster(remfrac, filename = full_path(p_bd, "ZA-remfrac.tif"))

# Calculate intactness
w <- matrix(nrow = 11, ncol = 11)
w[] <- 1
intact <- focal(remfrac, w = w, fun = mean, na.rm = TRUE)
writeRaster(intact, filename = full_path(p_bd, "intactness.tif"))

# intact <- mask(intact, mgrid)
# plot(intact)

# Calculate a veg type weight based on relative remaining area
veg_dt <- dtraster::as.data.table(sm)  # stack and convert to dt
veg_dt <- na.omit(veg_dt)
vnms <- names(veg_dt)[-c(1:2)]
vtots <- veg_dt[, lapply(.SD, function(x) sum(x * 100)), 
                .SDcols = vnms]  
1 / wgti
wgti <- round(1 / (vtots / nrow(veg_dt)), 2)  # bd weights (1 = rarest)
# wgt <- round(vtots / nrow(veg_dt), 4)  # biodiversity weights (1 = most common)
# plot(unlist(wgti), unlist(wgt))

@@

# plots to examine them
labs <- gsub("veg", "", names(wgtsn))
# par(mfrow = c(2, 1), mar = c(2, 2, 2, 2))
# weights against protected proportion (NPs + GMAs)
plot(wgtsn, prop_prot2, pch = "", ylab = "")
text(wgtsn, prop_prot2, labels = labs, cex = 0.75)
# weights against protected proportion (NPs only)
plot(wgtsn, prop_prot1, pch = "", ylab = "")
text(wgtsn, prop_prot1, labels = labs, cex = 0.75)
# additive weights factoring in protected proportion versus protected proportion
plot(wgts_prop2, prop_prot, pch = "", ylab = "")
text(wgts_prop2,  prop_prot, labels = labs, cex = 0.75)
# BD + NP protected proportion versus NP protected proportion
plot(wgts_prop_np, prop_prot1, pch = "", ylab = "")
text(wgts_prop_np,  prop_prot1, labels = labs, cex = 0.75)
# protected proportion weighted versus same but log-transformed
plot(wgts_prop, wgts_propl, pch = "", ylab = "")
text(wgts_prop,  wgts_propl, labels = labs, cex = 0.75)
# Bd weights against BD + protected proportion weights
plot(wgtsn, wgts_prop, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn, wgts_prop, labels = labs, cex = 0.75)
# Bd weights against BD + NP protected proportion weights
plot(wgtsn, wgts_prop_np, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn, wgts_prop_np, labels = labs, cex = 0.75)
# Bd weights versus log-transformed BD + protected proportion weights
plot(wgtsn, wgts_propl, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn,  wgts_propl, labels = labs, cex = 0.75)
# The same but against multiplicative weights
plot(wgtsn, wgtsn * unlist(1 - prop_prot1), pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn,  wgtsn * unlist(1 - prop_prot1), labels = labs, cex = 0.75)
# sort(wgts_prop); sort(wgts_prop_np)
# sort(wgtsn)

@@

# final biodiversity layer
bdf <- calc(stack(bdr, pafix), max)
# plot(remfrac)
# plot(bdf)

writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"), overwrite = TRUE)

vprops <- cbind("var" = c("area", "rarity", "priority"),
                rbind(vtots / nrow(veg_dt), wgti, data.table(t(wgts_prop_np))))
save(vprops, file = full_path(p_bd, "vprops.rda"))

### 2016-01-19 Rmd/cropland.Rmd
Updated cropland datasets, calculating current productivity for maize and soybean from FAO and USDA statistics.

---
title: "cropland.Rmd"
author: "Lyndon Estes"
date: "January 14, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

@@

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_lc <- "~/Dropbox/data/landcover/Zambia_LandCover_2010_Scheme_II/"

@@

onm <- full_path(p_devel, "cropland-1k.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = cropland@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
r <- raster(onm)
par(mar = rep(0, 4)); plot(r)

@@

>>>>

---
title: "Current Cropland, Production, and Yield"
author: "Lyndon Estes"
date: "January 14, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

@@

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_dat <- "~/Dropbox/data/agricultural/fao/countrystat/zambia/"
p_usda <- "~/Dropbox/data/agricultural/usda/fas/zambia/"
p_lc <- "~/Dropbox/data/landcover/Zambia_LandCover_2010_Scheme_II/"

@@

onm <- full_path(p_devel, "cropland-1k.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = cropland@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
cropland <- raster(onm)
par(mar = rep(0, 4)); plot(cropland)

@@

## Cropland statistics

From FAO countrystat at this [link](http://www.countrystat.org/home.aspx?c=ZMB&tr=21).
```{r, eval = FALSE}
# select crop stats
fnms <- dir(p_dat, pattern = "10|16", full.names = TRUE)
vars <- c("prod", "pa")
fao <- rbindlist(lapply(1:length(fnms), function(x){
  DT <- fread(fnms[x], stringsAsFactors = FALSE)
  cbind(vars[x], DT)
})) 
fao[, Value := as.numeric(Value)]
setnames(fao, old = c("V1", "administrative level 1 (code)", 
                      "administrative level 1", "Crop (code)"), 
         new = c("var", "provcode", "prov", "cropcode"))
setkeyv(fao, c("var", "Year"))

# missing 2008 NW province maize and 2011 Southern Province, same for soybean
fao[Crop == "Maize" & var == "prod" & is.na(Value)]  
fao[Crop == "Soybeans" & var == "prod" & is.na(Value)]

# corresponding FAOStat full country values are 1211566 and 3020380 for maize
# 56839 and 116539 for soybean
nwrepl <- fao[Crop == "Maize" & var == "prod" & Year == "2008" & 
                provcode != 3431, 1211566 - sum(Value)]
fao[Crop == "Maize" & var == "prod" & Year == "2008" & 
                provcode == 3431, Value := nwrepl]
srepl <- fao[Crop == "Maize" & var == "prod" & Year == "2011" & 
               provcode != 3433, 3020380 - sum(Value)]
fao[Crop == "Maize" & var == "prod" & Year == "2011" & 
                provcode == 3433, Value := srepl]
nwrepl <- fao[Crop == "Soybeans" & var == "prod" & Year == "2008" & 
                provcode != 3431, 56839 - sum(Value)]
fao[Crop == "Soybeans" & var == "prod" & Year == "2008" & 
                provcode == 3431, Value := nwrepl]
srepl <- fao[Crop == "Soybeans" & var == "prod" & Year == "2011" & 
               provcode != 3433, 116539 - sum(Value)]
fao[Crop == "Soybeans" & var == "prod" & Year == "2011" & 
                provcode == 3433, Value := srepl]

# Planted area and production 2010-2014 for Zambia
muprod <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "prod" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(muprod) <- c("maize", "soy")
mupa <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "pa" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(mupa) <- c("maize", "soy")

# Read in USDA harvested area and production to get difference in harvested area
# and planted area
fnms <- dir(p_usda, full.names = TRUE)
vars <- c("maize", "soy")
usda <- rbindlist(lapply(1:length(fnms), function(x){
  DT <- fread(fnms[x], stringsAsFactors = FALSE, header = TRUE)
  # cbind(vars[x], DT)
}))
cols <- names(usda)[4:9]
usda_ylds <- usda[var == "prod", cols, with = FALSE] / 
  usda[var == "harvarea", cols, with = FALSE]

# FAO references start year, it seems for maize, but soybean is correct year
yp <- unlist(usda[var == "prod" & crop == "maize", cols, with = FALSE] * 1000)
plot(yp, muprod[[1]]$V1[-c(1, 8)])  # but essentially same yields
yp <- unlist(usda[var == "prod" & crop == "soybean", cols, with =FALSE] * 1000)
plot(yp, muprod[[2]]$V1[-c(1:2)])

# Production estimates ~2010/1-2015
currprod <- round(c("maize" = muprod$maize[Year %in% 2010:2014, mean(V1)], 
                    "soy" = muprod$soy[Year %in% 2011:2015, mean(V1)]))
save(currprod, file = full_path(p_devel, "currprod.rda"))

## Resulting model inputs

`currprod.rda`, `cropland-1k.tif`

Eventually might want to disaggregate yields onto cropland map, but will need provincial harvested area for that.  Model might not really need that, however.
